---
title: "Deep Learning Excursion 1: ISLR's CNN for classifying a subset of CIFAR100 images"
author: "Vincent J. Carey, stvjc at channing.harvard.edu"
date: "`r format(Sys.time(), '%B %d, %Y')`"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Deep Learning excursion 1: ISLR's CNN for classifying a subset of CIFAR100 images}
  %\VignetteEncoding{UTF-8}
output:
  BiocStyle::html_document:
    highlight: pygments
    number_sections: yes
    theme: united
    toc: yes
---

# Introduction

The CIFAR100 data includes some challenging image categories,
and the demonstration convolutional neural network (CNN) examined
in the ISLR book has relatively low accuracy over the entire
set of images.  In this vignette we will use the ImageArray
class operations to form a subset of images of boys, roses, and chairs,
train with these images, and assess accuracy.

# Build ImageArray instances

We need two packages:
```{r makeima,message=FALSE}
library(littleDeep)
library(jsonlite)
```

We'll start by getting the fine-grained image types
from CIFAR100 JSON.

```{r dojs}
cifmeta = fromJSON(system.file("extdata", "cif.json", package="littleDeep"))
labs = cifmeta[[1]]$features$fine_label$names
head(labs)
```

Now get the image data and work with the training set
to produce an ImageArray.

```{r dotra}
if (!exists("cifdata")) cifdata = dataset_cifar100()
ciftrain = cifdata[[1]]$x
ciflabels = labs[cifdata[[1]]$y+1]

train = ImageArray(ciftrain, ciflabels)
train
```

Same for the test images:
```{r dotest}
ciftest = cifdata[[2]]$x
ciftestlabs = labs[cifdata[[2]]$y+1]
test = ImageArray(ciftest, ciftestlabs)
test
```

# Filter the ImageArrays to specific image classes

We make a vector of image types of interest and use `filterByType`
to subset the ImageArrays.
```{r dofilt}
littrain = filterByType(train, c("boy", "rose", "chair"))
littest = filterByType(test, c("boy", "rose", "chair"))
littrain
preview(littrain)
```

# Fit and assess

This can take a few minutes on a CPU.  We use the default
number of epochs.
```{r dofit}
rr = islr_cnn( littrain )
rr
rr$model
plot(rr$history)
```

To wrap up, we evaluate and have a look at predictions.

```{r fin}
eval_model(rr$model, littest)
save_islr_cnn(rr, "cif3")
model_probs(rr$model, littrain[1:9])
preview(littrain[1:9])
```
