---
title: "ISLR's CNN for classifying geometric shapes"
author: "Vincent J. Carey, stvjc at channing.harvard.edu"
date: "`r format(Sys.time(), '%B %d, %Y')`"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{ISLR's CNN for classifying geometric shapes"
  %\VignetteEncoding{UTF-8}
output:
  BiocStyle::html_document:
    highlight: pygments
    number_sections: yes
    theme: united
    toc: yes
---

# Introduction

In this vignette we consider how to classify images
representing geometric shapes.  The idea is that
we can produce very large training sets from specified
classes and explore relationships between complexity
of comparison tasks, training set size, and accuracy.

We'll work with the image-classifying CNN from ISLR.

# Building collections of randomly generated shapes

We have code that situates pixels
derived from randomly constructed and positioned circles, triangles, and
quadrilaterals in a 32 x 32 plane.  We form
ImageArray instances for each geometric type.
```{r setup, message=FALSE}
library(littleDeep)
library(abind)
set.seed(1234)
```


```{r buildthem}
NIMAGES=2500

build_array = function( shapefun = rancirc, nimages=NIMAGES, sidelength=64, depth=3, npts=90, ... ) {
 tmp = array(1, dim=c(nimages,sidelength,sidelength,depth)) # store images with 0 for dark
 for (i in seq_len(nimages)) {
    z = shapefun(npts=npts, ...)
    p = load_jpeg(z$x, z$y, siz=sidelength, dim=c(sidelength,sidelength,3))
    tmp[i,,,] = p
    }
  tmp
}
 

# training data

circarr = build_array( rancirc, sidelength=32, side_plane=32, npts=120 )
triarr = build_array( rantri, sidelength=32, side_plane=32, npts=40 )
quadarr = build_array( ranquad, sidelength=32, side_plane=32, npts=30 )

allsh = abind(circarr, triarr, along=1)
allsh = abind(allsh, quadarr, along=1)

shapes = ImageArray(allsh, types=rep(c("circle", "triangle", "quad"), each=NIMAGES))

# test data

test_circarr = build_array( rancirc, sidelength=32, side_plane=32, npts=120 )
test_triarr = build_array( rantri, sidelength=32, side_plane=32, npts=40 )
test_quadarr = build_array( ranquad, sidelength=32, side_plane=32, npts=30 )

test_allsh = abind(test_circarr, test_triarr, along=1)
test_allsh = abind(test_allsh, test_quadarr, along=1)

test_shapes = ImageArray(test_allsh, types=rep(c("circle", "triangle", "quad"), each=NIMAGES))
preview(shapes[c(1:3, 2501:2503, 5001:5003)])
```

# Fit the ISLR CNN on the training data

```{r lkcnn}
fit_sh = islr_cnn( shapes )
plot(fit_sh$history)
eval_model(fit_sh$model, test_shapes)
```
